---
title: "digits-recognition"
author: "Hugh"
date: "29 November 2016"
output: html_document
---

```{r}
digits <- read.csv("../data-raw/digits/train.csv")

# Two ways to determine the time taken in an short line of code:
if (FALSE){  # if (FALSE): don't run (too slow to do each time)
system.time(digits <- read.csv("../data-raw/digits/train.csv"))

# times = 5L means repeat the expression five times
# to obtain an average.
library(microbenchmark)
microbenchmark(digits <- read.csv("../data-raw/digits/train.csv"), times = 5L)
}

# What's a faster way to read in data?
library(data.table)
library(readr)
system.time(digits <- read_csv("../data-raw/digits/train.csv"))
system.time(digits <- fread("../data-raw/digits/train.csv"))
```

What are we looking at?
```{r}
head(digits)
dim(digits)
library(dplyr)
glimpse(digits)
```

This data contains 784 columns representing the darkness of the pixel in a 28 by 28 image and `label` the digit that was actually drawn.

## Visualizing the data

For a random observation, plot an image of what the observation represents: i.e. what was actually drawn by the person. (Use `ggplot2`.) Also plot in the bottom-right corner an annotation of the actual label. 









