---
title: "digits-recognition"
author: "Hugh"
date: "29 November 2016"
output: html_document
---

```{r}
digits <- read.csv("../data-raw/digits/train.csv")

# Two ways to determine the time taken in an short line of code:
if (FALSE){  # if (FALSE): don't run (too slow to do each time)
system.time(digits <- read.csv("../data-raw/digits/train.csv"))

# times = 5L means repeat the expression five times
# to obtain an average.
library(microbenchmark)
microbenchmark(digits <- read.csv("../data-raw/digits/train.csv"), times = 5L)
}

# What's a faster way to read in data?
library(data.table)
library(readr)
system.time(digits <- read_csv("../data-raw/digits/train.csv"))
system.time(digits <- fread("../data-raw/digits/train.csv"))
```

What are we looking at?
```{r}
head(digits)
dim(digits)
library(dplyr)
glimpse(digits)
```

This data contains 784 columns representing the darkness of the pixel in a 28 by 28 image and `label` the digit that was actually drawn.

## Visualizing the data

For a random observation, plot an image of what the observation represents: i.e. what was actually drawn by the person. (Use `ggplot2`.) Also plot in the bottom-right corner an annotation of the actual label.

```{r}
digits %>%
  sample_n(1) %>%
  select(-1) %>%
  matrix(nrow = 28, ncol = 28)
```


```{r}
library(magrittr)
library(ggplot2)
data.table::CJ(x = 1:2, y = 1:2) %>%
  mutate(grey = runif(n())) %>%
  ggplot + 
  geom_tile(aes(x = x, y = y, alpha = grey)) 
```

In order to plot a row, we need to get the x coordinates, the y coordinates, and the pixel transparency at each of them, in **separate columns of a single data frame**.

To turn the digits data from wide to long form, we need to use `melt` from `data.table` package, or `gather` from the tidyr package.

```{r, eval = FALSE}
library(tidyr)
digits %>%
  gather(pixel, greyness, -label)
```

```{r}
library(dtplyr)
# digits <- fread(...)
digits %>%
  sample_n(1) %>%
  melt.data.table(id.vars = "label") %>%
  # We want to remove 'pixel' from the 'variable' column, and put
  # this into a new column
  mutate(pixel_no = as.numeric(gsub("pixel", "", variable))) %>%
  # Assume the digits go up columns
  mutate(y_coord = floor(pixel_no / 28)) %>% 
  mutate(x_coord = pixel_no %% 28) %>% 
  ggplot + 
  geom_tile(aes(x = x_coord, y = -y_coord, alpha = value)) +
  geom_text(x = 28, y = -28, aes(label = label))
```







